\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2022-03-03 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

\begin{figure*}[ht!]
   \subfigure[Case I]{\includegraphics[width=0.49\textwidth]{IEEE14_Case0_RotorAngles.png}}
   \hspace{\fill}
   \subfigure[Case II]{\includegraphics[width=0.49\textwidth]{IEEE14_Case1_RotorAngles.png}}
\caption{\small{Boxplots for the rotor angles $\theta_{i}(t)$, $i=1,\hdots,5$, for the IEEE 14 bus system simulation in Sec. \ref{subsec:NumSimIEEEBus} in the (a) nominal \textbf{Case I}, and in the post-fault \textbf{Case II} scenarios, respectively. As usual, the boxes show the respective interquartile ranges, and the whiskers show the respective minimum and maximum values. In each box plot, we have superimposed  the respective \emph{mean} rotor angle sample path.}}
\vspace*{-0.1in}
\label{fig:IEEE14theta}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic50Gen.png}
\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{50}\times\mathbb{R}^{50}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.1in}
\label{fig:CompTimeSynthetic}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic50Gen.png}
\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.2in}
\label{fig:RelErrSynthetic}
\end{figure}

@@

To depict the numerical accuracy, Fig. \ref{fig:RelErrSynthetic} plots the (time-varying) relative error between the empirical (i.e., Monte Carlo estimate) mean vector $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the ``proximal mean" vector $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$. The latter was computed using the proximal updates $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$. Fig. \ref{fig:RelErrSynthetic} shows that the time-varying statistics from these two computation match very well at all times.

@@

\bibitem{cui2021andes}
H.~Cui, F.~Li, and K.~Tomsovic, ``Hybrid Symbolic-Numeric Framework for Power System Modeling and Analysis,'' \emph{IEEE Transactions on Power Systems}, vol. 36, no. 2, pp. 1373-1384, 2021.

>>>>

\begin{figure*}[ht!]
   \subfigure[Case I]{\includegraphics[width=0.49\textwidth]{IEEE14_Case0_RotorAngles.png}}
   \hspace{\fill}
   \subfigure[Case II]{\includegraphics[width=0.49\textwidth]{IEEE14_Case1_RotorAngles.png}}
\caption{\small{Boxplots for the rotor angles $\theta_{i}(t)$, $i=1,\hdots,5$, for the IEEE 14 bus system simulation in Sec. \ref{subsec:NumSimIEEEBus} in the (a) nominal \textbf{Case I}, and in the (b) post-fault \textbf{Case II} scenarios, respectively. As usual, the boxes show the respective interquartile ranges, and the whiskers show the respective minimum and maximum values. In each box plot, we have superimposed  the respective \emph{mean} rotor angle sample path.}}
\vspace*{-0.1in}
\label{fig:IEEE14theta}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic50Gen.png}
\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{50}\times\mathbb{R}^{50}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.1in}
\label{fig:CompTimeSynthetic}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic50Gen.png}
\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.15in}
\label{fig:RelErrSynthetic}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{RelErrCovMCvsProxSythetic50Gen.png}
\caption{\small{The normalized Bures-Wasserstein distance $d_{\rm{BW}}(\cdot,\cdot)$ between the empirical covariance  (i.e., Monte Carlo estimate) ${\rm{Cov}}^{\rm{MC}}\in\mathbb{R}^{100\times 100}$, and the ensemble covariance ${\rm{Cov}}^{\rm{Prox}}\in\mathbb{R}^{100\times 100}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}. The normalization is w.r.t. $\sqrt{{\rm{trace}}\left({\rm{Cov}}^{\rm{MC}}\right)}$.}}
\vspace*{-0.2in}
\label{fig:RelErrCovSynthetic}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{WassMCvsProxSythetic50Gen.png}
\caption{\small{The 2-Wasserstein distance between the empirical joint PDF (i.e., Monte Carlo estimate) $\rho^{\rm{MC}}$, and the proximal update $\varrho^{k}$ obtained using the proposed proximal algorithm, for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.2in}
\label{fig:WassJointPDFSynthetic}
\end{figure}

@@

To depict the numerical accuracy, Fig. \ref{fig:RelErrSynthetic} plots the (time-varying) relative error between the empirical (i.e., Monte Carlo estimate) mean vector $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the ``proximal mean" vector $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$. The latter was computed using the proximal updates $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$. {\blue{Likewise, Fig. \ref{fig:RelErrCovSynthetic} plots the time-varying normalized Bures-Wasserstein distance $d_{\rm{BW}}(\cdot,\cdot)$ between the empirical covariance matrix (i.e., Monte Carlo estimate) ${\rm{Cov}}^{\rm{MC}}\in\mathbb{R}^{100\times 100}$, and the ensemble covariance matrix ${\rm{Cov}}^{\rm{Prox}}\in\mathbb{R}^{100\times 100}$ obtained using the proximal updates. The unnormalized Bures-Wasserstein distance \cite{bhatia2019bures} given by 
\begin{align*}
&d_{\rm{BW}}(\rm{Cov}^{\rm{MC}},\rm{Cov}^{\rm{Prox}}) := \left[\operatorname{trace}\left(\rm{Cov}^{\rm{MC}}+\rm{Cov}^{\rm{Prox}}\right)\right.\\
&\qquad\qquad \left.-2\operatorname{trace}\left(\left(\sqrt{\rm{Cov}^{\rm{MC}}}\:\rm{Cov}^{\rm{Prox}}\:\sqrt{\rm{Cov}^{\rm{MC}}}\right)^{\frac{1}{2}}\right)\right]^{\frac{1}{2}},	
\end{align*}
is a metric on the cone of symmetric positive definite matrices such as the covariance matrices, and can be seen as an absolute error between the covariances. Thus, Fig. \ref{fig:RelErrCovSynthetic} can be viewed as a relative error curve between the covariances. In summary, Figs. \ref{fig:RelErrSynthetic} and \ref{fig:RelErrCovSynthetic} show that the time-varying statistics between the Monte Carlo and the proposed proximal computation remain close at all times.}}

{\blue{Since it is not possible to \emph{a priori} determine the dimension of the sufficient statistic of the transient joint state PDFs, Fig. \ref{fig:WassJointPDFSynthetic} shows the 2-Wasserstein metric between the empirical joint PDF (i.e., Monte Carlo estimate) $\rho^{\rm{MC}}$, and the proximal update $\varrho^{k}$. In other words, Fig. \ref{fig:WassJointPDFSynthetic} can be viewed as a nonparametric absolute error curve between $\rho^{\rm{MC}}$ and $\varrho^{k}$. Each snapshot value in Fig. \ref{fig:WassJointPDFSynthetic} is computed by solving the well-known optimal transport linear program (see e.g., \cite[Sec. 4.2-4.3]{halder2014probabilistic}, \cite[Ch. 3]{peyre2019computational}) using the corresponding time-varying weighted point clouds.}}

@@

\bibitem{cui2021andes}
H.~Cui, F.~Li, and K.~Tomsovic, ``Hybrid Symbolic-Numeric Framework for Power System Modeling and Analysis'', \emph{IEEE Transactions on Power Systems}, Vol. 36, No. 2, pp. 1373-1384, 2021.

% == Bures-Wasserstein ====

{\blue{\bibitem{bhatia2019bures}
R.~Bhatia, T.~Jain, and Y.~Lim, ``On the Bures--Wasserstein Distance between Positive Definite Matrices", \emph{Expositiones Mathematicae}, Vol. 37, No. 2, pp. 165--191, 2019.

\bibitem{halder2014probabilistic}
A.~Halder, and R.~Bhattacharya, ``Probabilistic Model Validation for Uncertain Nonlinear Systems", \emph{Automatica}, Vol. 50, No. 8, pp. 2038--2050, 2014.

\bibitem{peyre2019computational}
G.~Peyr{\'e}, and M.~Cuturi, ``Computational Optimal Transport: With Applications to Data Science", \emph{Foundations and Trends{\textregistered} in Machine Learning}, Vol. 11, No. 5-6, pp. 355--607, 2019.

}}
